\input{~/preamble.tex}

\usepackage{blindtext}

\usepackage{pgfplots}
\usepackage{subfiles}

\sectionfont{\color{blue}\selectfont}
\subsectionfont{\color{green}\selectfont}

\begin{document}
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{theorem*}{Theorem}
\newtheorem{definition}{Definition}[section]
{\LARGE \textbf{Unit 13: Series}}
\thispagestyle{empty}
\tableofcontents
\newpage
\clearpage
\setcounter{page}{1}
\section{Infinite sums: a cautionary tale}
\subfile{./sections/1}

\section{The definition of infinite sum}
\subfile{./sections/2}

\section{Telescopic series}
\subfile{./sections/3}

\section{Examples of divergent series from the definition}
\subfile{./sections/4}

\section{Geometric series}
\subfile{./sections/5}

\section{Linearity of series}
\subfile{./sections/6}

\section{The tail of a series}
\subfile{./sections/7}

\section{A necessary condition for convergence of series}
\subfile{./sections/8}

\section{Positive series}
\subfile{./sections/9}

\section{The integral test}
\subfile{./sections/10}

\section{Integral test examples}
\subfile{./sections/11}

\section{Comparison tests for series}
Comparison tests for series and improper integrals are the same. Results based on: \begin{itemize}
  \item A positive series may only be convergent or divergent to \(\infty\)
  \item To prove a positive series is convergent, we only need to prove it is not \(\infty\)
\end{itemize}

\thm{
  \textbf{\textsc{bct} for series} \\
  Let \(\sum_{n}^{\infty} a_n\) and \(\sum_{n}^{\infty} b_n\) be two series. \begin{itemize}
    \item Assume, for every \(n \in \N\), \(a \leq a_n \leq b_n\)
    \item \textsc{then} \begin{itemize}
            \item \textsc{if} \DS{\sum_{n}^{\infty} a_n = \infty}, \textsc{then} \DS{\sum_{n}^{\infty} b_n = \infty}.
            \item \textsc{if} \DS{\sum_{n}^{\infty} b_n < \infty}, \textsc{then} \DS{\sum_{n}^{\infty} a_n < \infty}.
          \end{itemize}
  \end{itemize}
}

\thm{
  \textbf{\textsc{lct} for series}\\
  Let \(\sum_{n}^{\infty} a_n\) and \(\sum_{n}^{\infty} b_n\) be two positive series. \begin{itemize}
    \item \textsc{if} the limit \DS{L = \lim_{n\to\infty} \frac{a_n}{b_n}} exists---is a number---and \(L > 0\).
    \item \textsc{then} \begin{align*}
            \sum_{n}^{\infty} a_n \quad \text{and} \quad \sum_{n}^{\infty} b_n
          \end{align*}
          are both convergent or both divergent.
  \end{itemize}
}
\newpage

\section{Alternating series}
\textbf{Definition:} A series \DS{\sum_{n}^{\infty} a_n} is alternating when \(\forall n, a_n a_{n+1} < 0\)\\ This means the terms ``alternate'' between positive and negative.

\textbf{Example:} \begin{align*}
  S   & = \sum_{n = 1}^{\infty} \frac{(-1)^{n+1}}{n} = 1 - \frac{1}{2} + \frac{1}{3} - \frac{1}{4} + \frac{1}{5} + \dots \\
  S_1 & = 1                                                                                                              \\
  S_2 & = 1 - \frac{1}{2}                                                                                                \\
  S_3 & = 1 - \frac{1}{2} + \frac{1}{3}                                                                                  \\
  S_4 & = 1 - \frac{1}{2} + \frac{1}{3} - \frac{1}{4}                                                                    \\
  \dots
\end{align*}
We notice from a graph that \DS{S_2 < S_4 < S_6 < \dots < S_5< S_3 < S_1}\\
\DS{\{ S_{2n} \}_n} is increasing and bounded above (by \(S_1\)).\\
\DS{\{S_{2n+1}\}} is decreasing and bounded below (by \(S_2\)). \\
Then by \textsc{mct}, both are convergent. \\
Call \DS{A = \lim_{n\to\infty} a_n}, \DS{B = \lim_{n\to\infty} b_{2n+1}}, then \\
\begin{align*}
  S_{2n+1}                    & = S_{2n} + \frac{1}{2n+1} \qquad \text{we can use the limit laws, since all three terms have limits} \\
  \lim_{n\to\infty}  S_{2n+1} & =\lim_{n\to\infty} S_{2n} +\lim_{n\to\infty} \frac{1}{2n+1}                                          \\
  B                           & = A + 0
\end{align*}

\thm{
\textbf{Lemma:}\\
Let \DS{\{c_n\}_{n}^{\infty}} be a sequence. \begin{itemize}
  \item \textsc{if} the sequences of even and odd terms \begin{align*}
          \{c_{2n}\}_{n}^{\infty} \quad \text{and} \quad \{c_{2n+1}\}_{n}^{\infty}
        \end{align*}
        are convergent to the same limit
  \item \textsc{then} the full sequence \DS{\{c_n\}_{n}^{\infty} } is also convergent to the same limit.
\end{itemize}
}

\thm{
  \textbf{Alternating series test}:\\
  Consider a series of the form \begin{align*}
    \sum_{n}^{\infty} (-1)^n b_n \quad \text{or} \quad \sum_{n}^{\infty} (-1)^{n+1} b_n
  \end{align*}
  \textsc{if} \begin{enumerate}
    \item \(\forall n, \quad b_n > 0\)
    \item the sequence \DS{\{b_n\}_{n}^{\infty} } is decreasing
    \item \DS{\lim_{n\to\infty} b_n = 0}
  \end{enumerate}
  \textsc{then} the series is convergent.
}

\newpage

\section{Estimating the value of an alternating series}
Estimate the value of \DS{S = \sum_{n = 1}^{\infty} \frac{(-1)^n}{n^4}} with an error smaller than \(0.001\).
First, we need to show that it is convergent---it satisfies the hypotheses of the alternating series theorem. By \textsc{ast}, the series is convergent. \\
The actual value is \DS{S = \sum_{n = 1}^{\infty} \frac{(-1)^n}{n^4} = \lim_{k\to\infty} S_k}. This suggests that we can use \DS{S_k} as an estimate. \\
Estimate: \DS{S_k = \sum_{n = 1}^{\infty} \frac{(-1)^n}{n^4}} for some large \(k\)? Which value of \(k\)? \\
We need the error to be smaller than a precise number. Error of estimation: \(\vert S - S_k \vert\).
\thm{
  \textbf{Alternating series theorem, part 2}:
  consider a series of the form \begin{align*}
    \sum_{n}^{\infty} (-1)^n b_n \quad \text{or} \quad \sum_{n}^{\infty} (-1)^{n+1} b_n
  \end{align*}
  \begin{itemize}
    \item \textsc{if} it satisfies the same three hypotheses as before
    \item \textsc{then} \DS{\vert S - S_k \vert < b_{k+1}}
  \end{itemize}
  where \(S_k\) is the \(k\)-th partial sum of the series.
}
Then, error of estimation: \(\vert S - S_k \vert < \frac{1}{(k+1)^4}\). We need to choose \(k\) so that \DS{\frac{1}{(k+1)^4} < 0.001}. Pick \(k\) to be \(5\). \\
Estimate: \DS{-1 + \frac{1}{2^4} - \frac{1}{3^4} + \frac{1}{4^4} - \frac{1}{5^4} \approx -0.94753...}.

\newpage

\section{Absolute convergence vs conditional convergence}
Is \DS{\sum_{n}^{\infty} \frac{\sin n}{n^2}} convergent? \\
What is the relation between the series \DS{\sum_{n}^{\infty} a_n} and \DS{\sum_{n}^{\infty} \vert a_n \vert}?

\thm{
  \textbf{Absolute convergence test} \\
  Let \DS{\sum_{n}^{\infty} a_n} be a series. \begin{itemize}
    \item \textsc{if} the series \DS{\sum_{n}^{\infty} \vert a_n\vert} is convergent
    \item \textsc{then} the series \DS{\sum_{n}^{\infty} a_n} is convergent
  \end{itemize}
}

We can look at \DS{\sum_{n}^{\infty} \frac{\vert \sin n \vert}{n^2}}. Since it is a positive series, we can use comparison tests. \\
\DS{0 \leq \frac{\vert \sin n \vert}{n^2} \leq \frac{1}{n^2}}. We know that \DS{\sum_{n}^{\infty} \frac{1}{n^2}} is convergent, so by \textsc{bct}, \DS{\sum_{n}^{\infty} \frac{\vert \sin n \vert}{n^2}} is also convergent. By \textsc{absolute convergence test}, \DS{\frac{\sin n}{n^2}} is also convergent.

\textbf{e.g.} Let \DS{p_n = \begin{cases}
    1 / n \quad \text{if \(n\) is prime} \\
    -1 / n \quad \text{otherwise}
  \end{cases}} \DS{\quad} is \DS{\sum_{n}^{\infty} p_n} convergent?

First, we look at \DS{\sum_{n}^{\infty} \vert p_n \vert}. \DS{\sum_{n}^{\infty} \vert p_n \vert = \sum_{n}^{\infty} \frac{1}{n} = \infty}. \\
The absolute convergence test does not apply, so we do not know if this series is convergent or divergent.

\df{
  A convergent series \DS{\sum_{n}^{\infty} a_n} is...\begin{itemize}
    \item \underline{absolutely convergent} when \DS{\sum_{n}^{\infty} \vert a_n \vert} is also convergent.
    \item \underline{conditionally convergent} when \DS{\sum_{n}^{\infty} \vert a_n \vert = \infty}.
  \end{itemize}
}

For instance, \DS{\sum_{n = 1}^{\infty} \frac{(-1)^{n+1}}{n}} is conditionally convergent. By the alternating series test, it is convergent. \DS{\sum_{n = 1}^{\infty} \frac{1}{n}}, its absolute value, however, is divergent. \\
For instance, \DS{\sum_{n = 1}^{\infty} \frac{(-1)^{n+1}}{n^2}} is absolutely convergent. By the alternating series test, it is convergent. \DS{\sum_{n = 1}^{\infty} \frac{1}{n^2}}, its absolute value, is also convergent.

\newpage
\section{Proof of the absolute convergence test}
\textbf{Notation:} ``\textsc{p.t}'' means ``positive terms'', ``\textsc{n.t}'' means ``negative terms''. For instance, \begin{align*}
  \sum_{n = 1}^{\infty} \frac{(-1)^{n+1}}{n} & = 1 - \frac{1}{2} + \frac{1}{3} - \frac{1}{4} + \frac{1}{5} - \frac{1}{6} + \frac{1}{7} + \dots \\
  \sum_{n = 1}^{\infty} \text{\textsc{p.t}}  & = 1 \qquad + \frac{1}{3} \qquad + \frac{1}{5} \qquad + \frac{1}{7} + \dots                      \\
\end{align*}
Essentially, we are replacing the negative terms with zeros.
\begin{proof}~\\
  Assume \DS{\sum_{n}^{\infty} \vert a_n \vert < \infty}. \\
  Then, \DS{\sum_{n}^{\infty} \vert \text{\textsc{p.t}} \vert \leq \sum_{n}^{\infty} \vert a_n \vert} and \DS{\sum_{n}^{\infty} \vert \text{\textsc{n.t}} \vert \leq \sum_{n}^{\infty} \vert a_n \vert}. \\
  By \textsc{bct}, \DS{\sum_{n}^{\infty} \vert \text{\textsc{p.t}} \vert < \infty} and \DS{\sum_{n}^{\infty} \vert \text{\textsc{n.t}} \vert < \infty} \\
  Therefore, \DS{\sum_{n}^{\infty} \text{\textsc{p.t}}} and \DS{\sum_{n}^{\infty} \text{\textsc{n.t}}} are convergent.  \\
  By key observation \DS{\sum_{n}^{\infty} a_n = \sum_{n}^{\infty} \text{\textsc{p.t}} + \sum_{n}^{\infty} \text{\textsc{n.t}}}. Thus \DS{\sum_{n}^{\infty} a_n} is convergent.
\end{proof}



\newpage
\section{Infinite sums are not commutative}
This is at the heart of the difference between absolutely convergent and conditionally convergent series.
\begin{align*}
    1 - \frac{1}{2} + \frac{1}{3} - \frac{1}{4} + \frac{1}{5} - \frac{1}{6} + \frac{1}{7} - \frac{1}{8} + \frac{1}{9} - \frac{1}{10} + \frac{1}{11} - \frac{1}{12} + \ldots \\
    = \ln 2 \\
    1 + \frac{1}{3}- \frac{1}{2} +\frac{1}{5} + \frac{1}{7} - \frac{1}{4} + \frac{1}{9} + \frac{1}{11} - \frac{1}{6} + \frac{1}{13} + \frac{1}{15} - \frac{1}{8} + \ldots \\
    = \frac{3}{2} \ln 2
.\end{align*}
\thm{
    \begin{itemize}
        \item If a series is \emph{absolutely convergent}, then it can be ``reordered''. This won't change the value of the sum.
        \item If a series is \emph{conditionally convergent}, then 
            \begin{itemize}
                \item \ldots reordering it may change its value
                \item \ldots it can be reordered to make it convergent to any number we want
                \item \ldots it can be reordered to make it divergent in any way we want.
            \end{itemize}
    \end{itemize}
}

\newpage
\section{Ratio test}
To determine if a series is convergent or divergent, by computing a limit. \\
\fbox{
  \parbox{\textwidth}{
    Let \DS{\sum_{n}^{\infty} a_n} be a series. Assume \DS{\forall n, a_n \neq 0}. Assume the limit \begin{equation*}
      L= \lim_{n\to\infty} \left \vert \frac{a_{n_+1}}{a_n} \right \vert \quad \text{exists or is } \infty
    \end{equation*}
  }
}
\end{document}
